---
title: "Project One: Homelessness"
author: "Yuri Souza"
date: "2023-03-28"
output:
  pdf_document: default
  html_document: default
---

# **This file contains the workflow I used for the project topic Homelessness**

\pagebreak

# **First Version**

I choose to work with homeless students in New York state. I picked up this topic because this is probably the social class most affected by homelessness, given that they are young and must make decisions that will guide and impact adult life. The infographic is based on The Economist magazine style. Everything I used here is also available on my [*GitHub*](https://github.com/souzayuri/homeless_infographics_P1).

The infochartic was designed mainly in four sections. 
**1.** The first section includes a chart containing the total number of homeless students from 2009 to 2021. I decided to use a line chart because it is easier to see trends in time. Aligned with this chart, I included a map of students' homeless per county in New York state. I decided to use counties because I want to show that this problem is happening across the whole state. 

**2.**	Given that New York has 62 counties and it would be hard to make a chart representing each of them well, I tried to show these changes in the second section using the New York economic region's boundaries. I chose a stream chart to display the data over time because I wanted to show trends in the increased proportion between the areas simultaneously. Besides, it allowed me to avoid lines overlapping within the chart. I included a map showing the proportion increase in 2021 compared to 2009 per economic region. My idea of plotting both chart and map is to show that the rise in homeless students is unrelated to the region area size and that only looking at the percentage increase does not represent the number of students per region well. 

**3.**	In the third section, I use a line chart to show which schools these students attend and how disproportional it is over time. I also included a bar chart showing the proportion changes in 2021 compared to 2009. 

**4.** The fourth section is dedicated to the show the students' proportion per school grade over time, using a bumpchart. My choice for a bumpchat was because it allowed me to show the overlap between lines and trends better than the line chart. This section also has a bar chart comparing 2021 to 2009 for each grade. Since I do not have the data relative to students' age, I considered using grades to proxy how old these people living in homeless conditions are.

The figure below is a free-hand draft I made before designing the infographic using Adobe Illustrator®.
![](C:/Users/Yuri/My Drive/PhD/00_UM/03_graduate/00_courses/02_2nd_semester/02_infographic_I/01_class_recordings_and_slides/02_march/03-07-23/Project_1/homeless_infographics_P1/02_figure/project1_draft_YuriSouza.jpeg)
\pagebreak

The page below is the first version of the project proposed here. Some considerations I have for the second version:

**1.**	Maybe remove the bar chart of section three or include it within the line chart in this section since this bar chart seems a little redundant.

**2.**	Finding some neat way to replace or show the bumpchat chart in section four.

**3.**	Draw an image at the top of the last chart of section four. 

**4.**	Include information about where these students live when they are not in school.

\pagebreak

# **Second Version**

In the second version, I changed the second graph to a slope chart to avoid the repetition of chart aesthetics. Using a Sankey chart, I also included information about where these students live when they are not in school. I also tried filling the white/empty space with some information or aesthetics. 

\pagebreak

# **Third Version**

In the third version, I changed the third and fourth charts to a line and bar chart, respectively. The reason was to make the stylish less chaotic and to highlight the last chart, as it is the main chart on the infographic. I also removed the illustrations because they did not fit the story well and the magazine's style. I changed the colors according to The Economists requirements.

I am considering changing charts three and four about the homeless home after school for the fourth version since it seems to occupy considerable space and bring more attention than they should. For graph three, I am considering plotting the line chart above its respective bar graph, keeping them the same size. For graph four, I am thinking of splitting it into four charts (one for each type of home)

\pagebreak

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Scripts

## Loading the packages

```{r, echo=TRUE, results='hide', message = FALSE, include=TRUE}
if(!require(tidyverse))install.packages("tidyverse", dependencies = TRUE)
if(!require(readxl)) install.packages("readxl", dependencies = TRUE)
if(!require(purrr)) install.packages("purrr", dependencies = TRUE)
if(!require(writexl)) install.packages("writexl", dependencies = TRUE)
if(!require(openxlsx)) install.packages("openxlsx", dependencies = TRUE)
if(!require(data.table)) install.packages("data.table", dependencies = TRUE)
```

## Setting directories and reading tables

```{r, echo=TRUE, results='hide', message = FALSE, include=TRUE}

path <- "C:/Users/Yuri/My Drive/PhD/00_UM/03_graduate/00_courses/02_2nd_semester/02_infographic_I/01_class_recordings_and_slides/02_march/03-21-23/homeless_infographics_P1/00_data"
path

regions <- readr::read_csv("C:/Users/Yuri/My Drive/PhD/00_UM/03_graduate/00_courses/02_2nd_semester/02_infographic_I/01_class_recordings_and_slides/02_march/03-21-23/homeless_infographics_P1/00_data/regions_ny.csv") %>% 
  unique()
regions

file_path <- list.files(path, pattern="\\.xlsx$", full.names = TRUE)
file_path


#For each file in file_path
homeless.tables <- purrr::map(file_path, ~
                         #For each sheet
                         purrr::map(2:3, function(i) {
                           #Read the file with particular sheet nummber
                           openxlsx::read.xlsx(.x, sheet=i, startRow=1)}) %>% 
                         purrr::reduce(dplyr::full_join) %>%
                         #Remove all NA rows
                         #dplyr::filter(Reduce(`|`, across(.fns = ~!is.na(.)))) %>%
                         #Add Year column at 1st position
                         dplyr::mutate(YEAR = tools::file_path_sans_ext(basename(.x)), .before = 1))

homeless.tables
```

## Binding the tables in a single one

```{r, echo=TRUE, results='hide', message = FALSE, include=TRUE}
homeless.tables.bind <- data.table::rbindlist(homeless.tables, 
                                              use.names=TRUE, 
                                              fill=FALSE, 
                                              idcol=TRUE) %>% 
  dplyr::rename_all(., .funs = toupper) %>% 
  dplyr::filter(!"TOTAL.HOMELESS" == "s") %>% 
  dplyr::mutate(TOTAL.HOMELESS = as.numeric(TOTAL.HOMELESS),
                YEAR = as.numeric(stringr::str_extract(YEAR, ".0.."))) %>% 
  dplyr::left_join(regions) %>% 
  stats::na.omit() %>% 
  dplyr::select(c(3,4,6,5,2))

homeless.tables.bind

#write_csv(homeless.tables.bind, "C:/Users/Yuri/My Drive/PhD/00_UM/03_graduate/00_courses/02_2nd_semester/02_infographic_I/01_class_recordings_and_slides/02_march/03-07-23/Project_1/homeless_infographics_P1/00_data/homeless.tables.bind.csv")

```

## Summarizing the tables by total individuals per year

```{r, echo=TRUE, results='hide', message = FALSE, include=TRUE}

homeless.tables.bind.year <- homeless.tables.bind %>% 
  dplyr::group_by(YEAR) %>% 
  dplyr::summarise(TOTAL.HOMELESS = sum(TOTAL.HOMELESS))
homeless.tables.bind.year

#write_csv(homeless.tables.bind.year, "C:/Users/Yuri/My Drive/PhD/00_UM/03_graduate/00_courses/02_2nd_semester/02_infographic_I/01_class_recordings_and_slides/02_march/03-07-23/Project_1/homeless_infographics_P1/00_data/homeless.tables.bind.year.csv")

homeless.tables.bind.year.diff <- homeless.tables.bind.year %>% 
  dplyr::filter(YEAR == 2009 | YEAR ==  2021) %>% 
  tidyr::pivot_wider(names_from = YEAR, values_from = TOTAL.HOMELESS) %>% 
  dplyr::mutate(DIFF = `2021`-`2009`,
                PCT_INCREASE = ((DIFF*100)/(`2009`))) %>% 
  dplyr::filter(PCT_INCREASE != 0) %>% 
  dplyr::filter_all(all_vars(!is.infinite(.)))
homeless.tables.bind.year.diff

#write_csv(homeless.tables.bind.year.diff, "C:/Users/Yuri/My Drive/PhD/00_UM/03_graduate/00_courses/02_2nd_semester/02_infographic_I/01_class_recordings_and_slides/02_march/03-07-23/Project_1/homeless_infographics_P1/00_data/homeless.tables.bind.year.diff.csv")

```

## Summarizing the tables by total individuals per year and region

```{r, echo=TRUE, results='hide', message = FALSE, include=TRUE}

homeless.tables.bind.year.region <- homeless.tables.bind %>% 
  dplyr::group_by(REGION, YEAR) %>% 
  dplyr::summarise(TOTAL.HOMELESS = sum(TOTAL.HOMELESS)) %>% 
  dplyr::ungroup() 
homeless.tables.bind.year.region

#write_csv(homeless.tables.bind.year.region, "C:/Users/Yuri/My Drive/PhD/00_UM/03_graduate/00_courses/02_2nd_semester/02_infographic_I/01_class_recordings_and_slides/02_march/03-07-23/Project_1/homeless_infographics_P1/00_data/homeless.tables.bind.year.region.csv")

homeless.tables.bind.year.region.diff <- homeless.tables.bind.year.region %>% 
  dplyr::filter(YEAR == 2009 | YEAR ==  2021) %>% 
  tidyr::pivot_wider(names_from = YEAR, values_from = TOTAL.HOMELESS) %>% 
  dplyr::mutate(DIFF = `2021`-`2009`,
                PCT_INCREASE = ((DIFF*100)/(`2009`))) %>% 
  dplyr::filter(PCT_INCREASE != 0) %>% 
  dplyr::filter_all(all_vars(!is.infinite(.))) %>% 
  tidyr::pivot_longer(cols = 2:3, values_to = "INDI", names_to = "YEAR") %>% 
  dplyr::select(c(1,4,5,2,3))

homeless.tables.bind.year.region.diff

#write_csv(homeless.tables.bind.year.region.diff, "C:/Users/Yuri/My Drive/PhD/00_UM/03_graduate/00_courses/02_2nd_semester/02_infographic_I/01_class_recordings_and_slides/02_march/03-07-23/Project_1/homeless_infographics_P1/00_data/homeless.tables.bind.year.region.diff.csv")



homeless.tables.bind.year.region.pct.diff <- homeless.tables.bind.year.region %>% 
  #dplyr::filter(YEAR == 2009 | YEAR ==  2021) %>% 
  tidyr::pivot_wider(names_from = YEAR, values_from = TOTAL.HOMELESS) %>% 
  dplyr::mutate(`2011i` = (((`2009`-`2011`)*100)/(`2009`)),
                `2012i` = (((`2011`-`2012`)*100)/(`2011`)),
                `2013i` = (((`2012`-`2013`)*100)/(`2012`)),
                `2014i` = (((`2013`-`2014`)*100)/(`2013`)),
                `2015i` = (((`2014`-`2015`)*100)/(`2014`)),
                `2016i` = (((`2015`-`2016`)*100)/(`2015`)),
                `2017i` = (((`2016`-`2017`)*100)/(`2016`)),
                `2018i` = (((`2017`-`2018`)*100)/(`2017`)),
                `2019i` = (((`2018`-`2019`)*100)/(`2018`)),
                `2020i` = (((`2019`-`2020`)*100)/(`2019`)),
                `2021i` = (((`2020`-`2021`)*100)/(`2020`))) %>% 
  dplyr::filter_all(all_vars(!is.infinite(.))) #%>% 
  #tidyr::pivot_longer(cols = 2:13, values_to = "DIFF_PCT", names_to = "YEAR") %>% 
  #dplyr::select(c(1,3,2)) %>% 
  #mutate(DIFF_PCT = round(DIFF_PCT, digits = 0),
  #       DIFF_STS = ifelse(DIFF_PCT >= 0, "Positive", "Negative"))


homeless.tables.bind.year.region.pct.diff

#write_csv(homeless.tables.bind.year.region.pct.diff, "C:/Users/Yuri/My Drive/PhD/00_UM/03_graduate/00_courses/02_2nd_semester/02_infographic_I/01_class_recordings_and_slides/02_march/03-07-23/Project_1/homeless_infographics_P1/00_data/homeless.tables.bind.year.region.pct.diff.csv")



```

## Summarizing the tables by total individuals per year and county

```{r, echo=TRUE, results='hide', message = FALSE, include=TRUE}

homeless.tables.bind.year.county <- homeless.tables.bind %>% 
  dplyr::group_by(COUNTY, YEAR) %>% 
  dplyr::summarise(TOTAL.HOMELESS = sum(TOTAL.HOMELESS))
homeless.tables.bind.year.county

#write_csv(homeless.tables.bind.year.county, "C:/Users/Yuri/My Drive/PhD/00_UM/03_graduate/00_courses/02_2nd_semester/02_infographic_I/01_class_recordings_and_slides/02_march/03-07-23/Project_1/homeless_infographics_P1/00_data/homeless.tables.bind.year.county.csv")

homeless.tables.bind.year.county.diff <- homeless.tables.bind.year.county %>% 
  dplyr::filter(YEAR == 2009 | YEAR ==  2021) %>% 
  tidyr::pivot_wider(names_from = YEAR, values_from = TOTAL.HOMELESS) %>% 
  dplyr::mutate(DIFF = `2021`-`2009`,
                PCT_INCREASE = ((DIFF*100)/(`2009`))) %>% 
  dplyr::filter(PCT_INCREASE != 0) %>% 
  dplyr::filter_all(all_vars(!is.infinite(.))) %>% 
  tidyr::pivot_longer(cols = 2:3, values_to = "INDI", names_to = "YEAR") %>% 
  dplyr::select(c(1,4,5,2,3))
homeless.tables.bind.year.county.diff

#write_csv(homeless.tables.bind.year.county.diff, "C:/Users/Yuri/My Drive/PhD/00_UM/03_graduate/00_courses/02_2nd_semester/02_infographic_I/01_class_recordings_and_slides/02_march/03-07-23/Project_1/homeless_infographics_P1/00_data/homeless.tables.bind.year.county.diff.csv")


```

## Summarizing the tables by total individuals per year and school type

```{r, echo=TRUE, results='hide', message = FALSE, include=TRUE}

homeless.tables.bind.year.school <- homeless.tables.bind %>% 
  dplyr::group_by(LEA.TYPE, YEAR) %>% 
  dplyr::summarise(TOTAL.HOMELESS = sum(TOTAL.HOMELESS))
homeless.tables.bind.year.school

#write_csv(homeless.tables.bind.year.school, "C:/Users/Yuri/My Drive/PhD/00_UM/03_graduate/00_courses/02_2nd_semester/02_infographic_I/01_class_recordings_and_slides/02_march/03-07-23/Project_1/homeless_infographics_P1/00_data/homeless.tables.bind.year.school.csv")

homeless.tables.bind.year.school.diff <- homeless.tables.bind.year.school %>% 
  dplyr::filter(YEAR == 2009 | YEAR ==  2021) %>% 
  tidyr::pivot_wider(names_from = YEAR, values_from = TOTAL.HOMELESS) %>% 
  dplyr::mutate(DIFF = `2021`-`2009`,
                PCT_INCREASE = ((DIFF*100)/(`2009`))) %>% 
  dplyr::filter(PCT_INCREASE != 0) %>% 
  dplyr::filter_all(all_vars(!is.infinite(.))) %>% 
  tidyr::pivot_longer(cols = 2:3, values_to = "INDI", names_to = "YEAR") %>% 
  dplyr::select(c(1,4,5,2,3))
homeless.tables.bind.year.school.diff

#write_csv(homeless.tables.bind.year.school.diff, "C:/Users/Yuri/My Drive/PhD/00_UM/03_graduate/00_courses/02_2nd_semester/02_infographic_I/01_class_recordings_and_slides/02_march/03-07-23/Project_1/homeless_infographics_P1/00_data/homeless.tables.bind.year.school.diff.csv")

```
## Summarizing the tables by grades

```{r, echo=TRUE, results='hide', message = FALSE, include=TRUE}


homeless.ny_regions <- readr::read_csv("C:/Users/Yuri/My Drive/PhD/00_UM/03_graduate/00_courses/02_2nd_semester/02_infographic_I/01_class_recordings_and_slides/02_march/03-21-23/homeless_infographics_P1/00_data/regions_ny.csv") %>% 
  unique()

homeless.ny_grades_regions <- readr::read_csv("C:/Users/Yuri/My Drive/PhD/00_UM/03_graduate/00_courses/02_2nd_semester/02_infographic_I/01_class_recordings_and_slides/02_march/03-21-23/homeless_infographics_P1/00_data/00_homeless_student_grades.csv") %>%        
  dplyr::left_join(homeless.ny_regions) %>% 
  dplyr::select(c(20, 3:14,19)) %>% 
  tidyr::pivot_longer(cols = 2:13, values_to = "Grades", names_to = "Age") %>% 
  #readr::write_csv("00_data/ny_homeless_student_grades_region.csv") %>% 
  dplyr::group_by(Year, Age) %>% 
  summarise(value = sum(Grades)) %>% 
  dplyr::ungroup() %>% 
  dplyr::filter(Year == "2009" | Year == "2021") %>% 
  tidyr::pivot_wider(names_from = "Year", values_from = "value") %>% 
  dplyr::mutate(DIFF = `2021`-`2009`,
                PCT_INCREASE = ((DIFF*100)/(`2009`))) %>% 
  dplyr::filter(PCT_INCREASE != 0) #%>% 
  #readr::write_csv("00_data/ny_homeless_student_grades_region_diff.csv")
  


```

## Summarizing the tables per overnighting 

```{r, echo=TRUE, results='hide', message = FALSE, include=TRUE}


homeless.ny_regions <- readr::read_csv("C:/Users/Yuri/My Drive/PhD/00_UM/03_graduate/00_courses/02_2nd_semester/02_infographic_I/01_class_recordings_and_slides/02_march/03-21-23/homeless_infographics_P1/00_data/regions_ny.csv") %>% 
  unique()

homeless.ny_overnight_regions <- readr::read_csv("C:/Users/Yuri/My Drive/PhD/00_UM/03_graduate/00_courses/02_2nd_semester/02_infographic_I/01_class_recordings_and_slides/02_march/03-21-23/homeless_infographics_P1/00_data/homeless_student_home.csv") %>%        
  dplyr::left_join(homeless.ny_regions) %>% 
  dplyr::select(c(7, 2:6)) %>% 
  tidyr::pivot_longer(cols = 2:5, values_to = "Value", names_to = "Home") %>% 
  dplyr::filter(!Value == 0) %>% 
  dplyr::filter(Year == "2009" | Year == "2021") %>% 
  dplyr::mutate(Year_fct = as.character(Year)) %>% 
  readr::write_csv("C:/Users/Yuri/My Drive/PhD/00_UM/03_graduate/00_courses/02_2nd_semester/02_infographic_I/01_class_recordings_and_slides/02_march/03-21-23/homeless_infographics_P1/00_data/homeless.ny_overnight_regions.csv") 
  


```

\pagebreak

